% -------------------------------------------------------------------------
% WBS ID: RES-OPT-OBJ
% Deliverable: Objective Functions and Performance Measures
% Status: In progress
% Evidence status: Objective use deferred; hydrodynamic metrics identified
% Review status: Not reviewed
% Last updated: 2026-07-19
% Dependencies: RES-GEO-MET, RES-PHY-BAR, RES-PHY-LOD, RES-VER-MET
% Downstream interfaces: Future RES-OPT-SPEC
% -------------------------------------------------------------------------

\section{RES-OPT-OBJ --- Objective Functions and Performance Measures}
\label{sec:res-opt-obj}

\subsection{Purpose and Scope}
\label{sec:res-opt-obj-purpose}

This Deliverable records future objective-function candidates but does
not activate optimisation. In the current project phase, run-up,
overtopping, transmission/reflection, pressure, force, impulse, moment
and energy are comparison metrics for fixed-rigid barrier classes, not
optimisation objectives. Structural response, adequacy, damage
classification and surrogate uncertainty are deferred.

\subsection{Research Questions}
\label{sec:res-opt-obj-research-questions}

% TODO: State the questions that must be answered before the Deliverable
% can be accepted.

\subsection{Terminology and Definitions}
\label{sec:res-opt-obj-definitions}

% TODO: Define the technical terminology, symbols, quantities and
% classifications used in this Deliverable.

\subsection{Literature Evidence}
\label{sec:res-opt-obj-literature-evidence}

% TODO: Record evidence from peer-reviewed papers, authoritative datasets,
% standards, technical reports and primary documentation.
%
% Do not create a detached literature-summary list. Explain how each source
% supports, contradicts or limits a project decision.

\subsection{Governing Relations and Quantitative Definitions}
\label{sec:res-opt-obj-governing-relations}

% TODO: Record relevant equations, dimensionless groups, empirical models,
% extraction rules or quantitative definitions.
%
% Use align environments for displayed equations.
% Every displayed equation must have a unique \label{eq:...}.
% Do not place punctuation after displayed equations.

\subsection{Decision Variables}
\label{sec:res-opt-obj-decision-variables}

% TODO: Complete this RES-OPT Domain-specific subsection.

\subsection{Objective Functions}
\label{sec:res-opt-obj-objective-functions}

Future objective components may include reducing peak force,
overturning moment, overtopping, transmitted energy or downstream
hazard while preserving constructability. Displacement, stress,
permanent deformation, damage probability and surrogate uncertainty
are excluded until structural and ML extensions are activated and
validated.

\subsection{Constraints}
\label{sec:res-opt-obj-constraints}

% TODO: Complete this RES-OPT Domain-specific subsection.

\subsection{Sampling Strategy}
\label{sec:res-opt-obj-sampling-strategy}

% TODO: Complete this RES-OPT Domain-specific subsection.

\subsection{Optimisation Method}
\label{sec:res-opt-obj-optimisation-method}

% TODO: Complete this RES-OPT Domain-specific subsection.

\subsection{Computational Budget}
\label{sec:res-opt-obj-computational-budget}

% TODO: Complete this RES-OPT Domain-specific subsection.

\subsection{Termination and Selection Criteria}
\label{sec:res-opt-obj-termination-selection-criteria}

% TODO: Complete this RES-OPT Domain-specific subsection.

\subsection{Candidate Approaches}
\label{sec:res-opt-obj-candidate-approaches}

% TODO: Define the credible candidate approaches considered.

\subsection{Critical Comparison}
\label{sec:res-opt-obj-critical-comparison}

% TODO: Compare candidates using physically and project-relevant criteria.
% Do not select an approach solely on popularity or implementation ease.

\subsection{Assumptions and Applicability}
\label{sec:res-opt-obj-assumptions}

% TODO: State assumptions and the conditions under which they are valid.

\subsection{Limitations and Failure Conditions}
\label{sec:res-opt-obj-limitations}

% TODO: State where the evidence, model or selected approach ceases to be
% reliable or applicable.

\subsection{Project-Specific Conclusions}
\label{sec:res-opt-obj-conclusions}

For the active Studentship, these quantities are reported as
hydrodynamic comparison metrics only. No objective aggregation, final
design ranking or optimisation claim shall be made from unvalidated
outputs.

\subsection{Selected Approach and Rationale}
\label{sec:res-opt-obj-selected-approach}

The selected active approach is metric reporting, not optimisation.
Future objective functions may be defined after RES-VER-MET accepts
the relevant hydrodynamic quantities.

\subsection{Unresolved Decisions}
\label{sec:res-opt-obj-unresolved-decisions}

% TODO: Record unresolved questions, required evidence and decision owners.

\subsection{Requirements and Handoffs}
\label{sec:res-opt-obj-requirements}

% TODO: State requirements passed to other Research Domains, Software,
% verification activities or later execution work.

\subsection{Deliverable Completion Record}
\label{sec:res-opt-obj-completion-record}

% TODO: Record whether the Deliverable is:
%
% - Not started
% - In progress
% - Draft complete
% - Reviewed
% - Accepted
%
% Acceptance requires:
%
% - the research questions to be answered;
% - claims to be supported by appropriate evidence;
% - assumptions and limitations to be explicit;
% - the project-specific conclusion to be stated;
% - downstream requirements to be traceable;
% - unresolved decisions to be closed or formally deferred.
