\documentclass[12pt,a4paper]{article}

\usepackage[hidelayout]{../../../Latex/Style/coursework-report-core}

\addbibresource{res-ml.bib}

\title{RES-ML --- Machine-Learning Methodology}
\author{}
\date{\today}

\begin{document}

\maketitle

\tableofcontents
\clearpage

\section{Domain Purpose and Scope}
\label{sec:res-ml-domain-purpose}

RES-ML defines the machine-learning methodology for the revised research
scope. Under the 2026-07-19 scope decision, ML is deferred future
work. The active Studentship objective is to stabilise the 2011
Tohoku hydrodynamic pipeline, Kamaishi/Sendai corridor comparison and
fixed-rigid barrier-class metrics before any supervised-learning,
surrogate optimisation or generative geometry work begins.

The high-fidelity simulation workflow remains the authoritative source of
physical response. A future surrogate may be an assessment, screening
or sampling tool; it shall not be represented as replacing validated
fluid simulation, structural analysis or direct confirmation.

Machine-learning training is blocked until the selected tsunami event and the
relevant hydrodynamic models have passed the V1--V2--V3 validation
hierarchy recorded in the current scope decision. Historical or
reconstructed defence geometries are evidence about plausible cases,
not direct causal labels for optimal reinforcement; controlled,
validated simulations remain the only acceptable future training
targets.

\subsection{Active Scope}
\label{sec:res-ml-active-scope}

There is no active ML training scope in the current Research phase.
Future RES-ML work may cover:
\begin{itemize}
  \item response surrogates for validated hydrodynamic outputs;
  \item optional adequacy or damage classification after structural
    extensions exist;
  \item screening within a bounded barrier-class or intervention space;
  \item active selection of additional validated simulations;
  \item dimensionality reduction where correlated outputs or histories justify
    it;
  \item small-data algorithm assessment and non-ML baselines;
  \item predictive uncertainty and out-of-domain detection;
  \item direct-simulation fallback for uncertain or unsupported queries;
  \item integration with validated regional and local CFD workflows,
    and later structural/FSI workflows if activated.
\end{itemize}

\subsection{Deferred Scope}
\label{sec:res-ml-deferred-scope}

The following are deferred to later research and shall not be represented as
initial acceptance requirements:
\begin{itemize}
  \item unrestricted free-form geometry generation;
  \item surrogate optimisation during the active hydrodynamic phase;
  \item reinforcement learning;
  \item unrestricted inverse design;
  \item claims of cross-event generalisation;
  \item direct training from before/after geometry alone;
  \item replacement of high-fidelity confirmation.
\end{itemize}

\subsection{Supervised-Learning Relationship}
\label{sec:res-ml-supervised-learning-relationship}

The intended supervised-learning relationship is:
\begin{align}
  &\text{corridor descriptors}
  + \text{fixed-rigid geometry descriptors}
  + \text{validated hydrodynamic-condition descriptors}
  \notag\\
  &\qquad \rightarrow
  \text{hydrodynamic response}
  + \text{predictive uncertainty}
  \label{eq:res-ml-supervised-learning-relationship}
\end{align}

Decision role: \cref{eq:res-ml-supervised-learning-relationship} is a
future data relationship, not an active implementation requirement.
Structural response and damage labels are excluded until RES-STR
activates and validates those extensions.

\subsection{Research Gates}
\label{sec:res-ml-research-gates}

The ML workstream is governed by the following gates:
\begin{enumerate}
  \item \textbf{G1 --- Data-source feasibility.} Source data,
    CRS/datum, licence and preprocessing provenance are recorded. No
    ML training is permitted before this gate.
  \item \textbf{G2 --- Corridor definition.} Kamaishi and Sendai
    corridor geometry is locked. ML remains deferred.
  \item \textbf{G3 --- Regional replay readiness.} The Tohoku regional
    model and replay-boundary histories are ready for validation. No
    ML dataset may be accepted before this gate.
  \item \textbf{G4 --- Local hydrodynamic metric extraction.}
    Fixed-rigid local runs emit complete quality and metric labels.
    Future dataset design may begin, but model training remains
    blocked until validation status is known.
  \item \textbf{G5 --- V1--V2--V3 validation.} Solver capability,
    Tohoku hydrodynamics and local impact/loading evidence are
    accepted. Future ML training may be reconsidered after this gate.
  \item \textbf{G6 --- Barrier-class demonstration.} Barrier classes
    are compared with complete provenance. This is not surrogate
    validation.
\end{enumerate}

\section{Domain Deliverables}
\label{sec:res-ml-domain-deliverables}

\input{"RES-ML-ROL - Machine-Learning Role and Use-Case Definition/res-ml-rol.tex"}
\input{"RES-ML-DAT - Learning Dataset and Simulation-Data Requirements/res-ml-dat.tex"}
\input{"RES-ML-REP - Input, Output and Geometry Representation/res-ml-rep.tex"}
\input{"RES-ML-ALG - Learning Algorithm and Architecture Assessment/res-ml-alg.tex"}
\input{"RES-ML-OBJ - Loss Functions, Rewards and Physical Constraints/res-ml-obj.tex"}
\input{"RES-ML-SUR - Surrogate, Multifidelity and Active-Learning Strategy/res-ml-sur.tex"}
\input{"RES-ML-TRN - Training, Validation and Experiment Methodology/res-ml-trn.tex"}
\input{"RES-ML-GEN - Generalisation, Uncertainty and Interpretability/res-ml-gen.tex"}
\input{"RES-ML-INT - Integration with Simulation and Optimisation/res-ml-int.tex"}
\input{"RES-ML-SPEC - Integrated Machine-Learning Specification/res-ml-spec.tex"}

\clearpage
\printbibliography[
  heading=bibintoc,
  title={References}
]

\end{document}
