\documentclass[12pt,a4paper]{article}

\usepackage[hidelayout]{../../../Latex/Style/coursework-report-core}

\addbibresource{res-dat.bib}

\title{RES-DAT --- Environmental Data and Case-Study Definition}
\author{}
\date{\today}

\begin{document}

\maketitle

\tableofcontents
\clearpage

\section{Domain Purpose and Scope}
\label{sec:res-dat-domain-purpose}

RES-DAT defines the environmental data and case-study evidence needed to
reconstruct one selected historical tsunami event, validate that event against
authoritative observations, and support documented defence-specific impact
cases. The initial project retains one primary historical event and controlled
physical variations around that event. It does not add unrelated tsunami events
or build a broad multi-event ML dataset during the initial scope.

The active studentship case is the 2011 Tōhoku event with two
regional-to-local corridor trajectories: epicentre to the Kamaishi
region and epicentre to the Sendai coastal plain. The active use of the
data is impact modelling, Tōhoku/local comparison and barrier-class
comparison. ML datasets, surrogate optimisation and generative geometry
remain future work until the event reconstruction, local hydrodynamic
load extraction and validation hierarchy are stable.

The Domain shall also catalogue defence geometry and local exposure
evidence where available, but damage, material deformation, scour and
FSI remain stretch extensions after hydrodynamic load extraction is
stable. Source role: project-scope baseline and data-interface
decision. Supporting sources: event metadata from
\textcite{usgs2011tohokuEvent}, onshore observations from
\textcite{mori2012nationwide}, offshore observations from
\textcite{noaa2005dart} and \textcite{noaa2011japanDart}, and terrain
source candidates in \textcite{jodcJegg500}, \textcite{gebco2026grid},
\textcite{noaa2022etopo} and \textcite{gsiDemDownload}. Limitation:
the cited source records support data selection and handoff
requirements; they do not demonstrate that the project has executed
validation.

\section{Domain Deliverables}
\label{sec:res-dat-domain-deliverables}

\input{"RES-DAT-GEO - Bathymetry, Topography and Coastal Geometry/res-dat-geo.tex"}
\input{"RES-DAT-SRC - Earthquake and Tsunami Source Data/res-dat-src.tex"}
\input{"RES-DAT-OBS - Observational and Validation Data/res-dat-obs.tex"}
\input{"RES-DAT-SCN - Scenario and Severity Definition/res-dat-scn.tex"}
\input{"RES-DAT-UNC - Data Quality, Uncertainty and Provenance/res-dat-unc.tex"}
\input{"RES-DAT-CAS - Integrated Tsunami Case-Study Specifications/res-dat-cas.tex"}

\clearpage
\printbibliography[
  heading=bibintoc,
  title={References}
]

\end{document}
