% -------------------------------------------------------------------------
% WBS ID: RES-GEO-REP
% Deliverable: Two- and Three-Dimensional Model Representation
% Status: In progress
% Evidence status: Fixed-rigid local representation established
% Review status: Not reviewed
% Last updated: 2026-07-19
% Dependencies: RES-GEO-PAR, RES-GEO-MET, RES-NUM-SPEC
% Downstream interfaces: RES-MOD-IBC, RES-NUM-MSH, RES-STR-LOD, RES-VER-SPEC, Software mesh/field containers
% -------------------------------------------------------------------------

\section{RES-GEO-REP --- Two- and Three-Dimensional Model Representation}
\label{sec:res-geo-rep}

\subsection{Purpose and Scope}
\label{sec:res-geo-rep-purpose}

This Deliverable defines how each active fixed-rigid barrier geometry
is represented for the regional model, local 3D fluid model,
hydrodynamic-load handoff and later deferred structural analysis. ML
feature generation is not active; any feature representation is a
future consumer of the same manifest.

\subsection{Research Questions}
\label{sec:res-geo-rep-research-questions}

\begin{enumerate}
  \item What is the minimum regional footprint required for regional
    propagation and replay extraction?
  \item What local 3D surface representation is required for
    no-slip/no-penetration wall patches and load extraction?
  \item Which representation fields must be preserved for later
    structural or ML use without activating those workstreams now?
\end{enumerate}

\subsection{Terminology and Definitions}
\label{sec:res-geo-rep-definitions}

% TODO: Define the technical terminology, symbols, quantities and
% classifications used in this Deliverable.

\subsection{Literature Evidence}
\label{sec:res-geo-rep-literature-evidence}

\textcite{watanabe2022validation} supports explicit local geometry,
probe and loading-output control in tsunami-impact validation.
\textcite{deljesus2012porous} supports distinguishing resolved solid
geometry from unresolved porous closure. Project conclusion: the
local representation shall be a resolved fixed solid surface unless a
separate porous model is later accepted.

\subsection{Governing Relations and Quantitative Definitions}
\label{sec:res-geo-rep-governing-relations}

The fixed barrier patch in the local fluid domain is:

\begin{align}
  \Gamma_{\mathrm{B}}
  &=
  \partial\Omega_{\mathrm{f}}
  \cap
  \partial\Omega_{\mathrm{B}}
  \label{eq:res-geo-rep-fixed-barrier-patch}
\end{align}

where $\Omega_{\mathrm{f}}$ is the local fluid-control domain and
$\Omega_{\mathrm{B}}$ is the solid barrier volume. Decision role:
\cref{eq:res-geo-rep-fixed-barrier-patch} defines the geometric
surface on which no-slip/no-penetration conditions and hydrodynamic
load integration are applied. Supporting evidence:
\textcite{watanabe2022validation}. Limitation: it does not describe
deformable FSI interfaces.

\subsection{Geometry Family or Representation}
\label{sec:res-geo-rep-geometry-family-representation}

% TODO: Complete this RES-GEO Domain-specific subsection.

\subsection{Design Variables and Parameter Bounds}
\label{sec:res-geo-rep-design-variables-parameter-bounds}

% TODO: Complete this RES-GEO Domain-specific subsection.

\subsection{Geometric Descriptors}
\label{sec:res-geo-rep-geometric-descriptors}

% TODO: Complete this RES-GEO Domain-specific subsection.

\subsection{Placement and Arrangement Rules}
\label{sec:res-geo-rep-placement-arrangement-rules}

% TODO: Complete this RES-GEO Domain-specific subsection.

\subsection{Feasibility Constraints}
\label{sec:res-geo-rep-feasibility-constraints}

% TODO: Complete this RES-GEO Domain-specific subsection.

\subsection{Two- and Three-Dimensional Representation}
\label{sec:res-geo-rep-two-three-dimensional-representation}

Representation levels shall include:

\begin{itemize}
  \item a regional-scale footprint or roughness/obstruction marker,
    used only if the regional model needs a defence representation;
  \item a local three-dimensional fixed solid surface for the
    URANS--VOF barrier interaction model;
  \item a patch and face identifier set for pressure, shear, force,
    impulse and moment extraction;
  \item a deferred structural surface mapping for one-way CFD-to-FEA
    load transfer after hydrodynamic extraction is stable;
  \item optional future ML descriptors generated from the same
    manifest, not an independent geometry source.
\end{itemize}

\subsection{Candidate Approaches}
\label{sec:res-geo-rep-candidate-approaches}

% TODO: Define the credible candidate approaches considered.

\subsection{Critical Comparison}
\label{sec:res-geo-rep-critical-comparison}

% TODO: Compare candidates using physically and project-relevant criteria.
% Do not select an approach solely on popularity or implementation ease.

\subsection{Assumptions and Applicability}
\label{sec:res-geo-rep-assumptions}

% TODO: State assumptions and the conditions under which they are valid.

\subsection{Limitations and Failure Conditions}
\label{sec:res-geo-rep-limitations}

% TODO: State where the evidence, model or selected approach ceases to be
% reliable or applicable.

\subsection{Project-Specific Conclusions}
\label{sec:res-geo-rep-conclusions}

The local solver needs an explicit fixed surface representation before
any barrier comparison is meaningful. The same surface identifiers
must survive into post-processing so forces and moments can be traced
back to geometry cases.

\subsection{Selected Approach and Rationale}
\label{sec:res-geo-rep-selected-approach}

The selected local representation is a fixed rigid solid volume
converted to terrain/barrier wall patches and stable patch identifiers.
Moving boundaries, topology change and deformable surfaces are
deferred to RES-STR-FSI.

\subsection{Unresolved Decisions}
\label{sec:res-geo-rep-unresolved-decisions}

Open decisions are the exact geometry file format, wall roughness
metadata, CAD/CSG versus raster-derived surface generation, and
Kamaishi real-defence source geometry.

\subsection{Requirements and Handoffs}
\label{sec:res-geo-rep-requirements}

Software shall support fixed geometry ingestion, patch labelling,
face/patch metric extraction and structured export to HDF5 or an
equivalent schema. RES-STR-LOD shall consume the same patch identifiers
for later hydrodynamic load transfer.

\subsection{Deliverable Completion Record}
\label{sec:res-geo-rep-completion-record}

% TODO: Record whether the Deliverable is:
%
% - Not started
% - In progress
% - Draft complete
% - Reviewed
% - Accepted
%
% Acceptance requires:
%
% - the research questions to be answered;
% - claims to be supported by appropriate evidence;
% - assumptions and limitations to be explicit;
% - the project-specific conclusion to be stated;
% - downstream requirements to be traceable;
% - unresolved decisions to be closed or formally deferred.
